---
title: "MT ANOVAS"
author: "Summer Clay"
date: "Monday, September 07, 2015"
output: word_document
---

### Model Overview
* Dependent Variable(s)
    + Reaction Time   
* Independent Variable(s)
    + Accept (Accept, Reject)
    + Card Number (1, 3, 5)
    + Deck Type (Win, Loss)
    + Block (1, 2, 3)
    + Confidence (Weak, Strong)

### Data Wrangling
  + Read in text file `DDMtask.txt` to dataframe `DDMtask`
  + Rename variable according to R style guide
  + Turn `subject`, `trial`, `accept`, `card.number`, `deck.type`, `cond.block`, `confidence` variables into R factors
  + Organize missing values into `NA` factor level
```{r, echo = FALSE}
# Read in data
DDMtask <- read.delim("C:/Users/Summer/Desktop/DDMtask.txt")
# Stylize names
names(DDMtask) <- c("subject", "trial", "accept", "rt", 
                    "card.number", "deck.type", "block", 
                    "gamble.outcome", "gamble.win", "confidence")
```

```{r, echo = FALSE}
# Transform 'accept', 'card.number', deck.type', 
# 'cond.block', 'confidence' as factors
DDMtask$subject <- as.factor(DDMtask$subject)
DDMtask$accept <- as.factor(DDMtask$accept)
DDMtask$card.number <- as.factor(DDMtask$card.number)
DDMtask$deck.type <- as.factor(DDMtask$deck.type)
DDMtask$confidence <- as.factor(DDMtask$confidence)
# Streamline missing values into 'NA'
levels(DDMtask$accept) <- c("Reject", "Accept", NA)
levels(DDMtask$deck.type) <- c("Loss", "Win")
levels(DDMtask$block) <- c(1, 2, 3, 1, 2, 3)
levels(DDMtask$confidence) <- c("Weak", "Strong", NA)
# remove cases where rt < 200ms
DDMtask.clean <- subset(DDMtask, rt >= 200)
```

### Descriptives: Reaction Time (Var 1) vs. logged Reaction Time (Var 2)
```{r, echo = FALSE, results = "asis", message = FALSE}
# log reaction time
DDMtask.clean$log.rt <- log(DDMtask.clean$rt)
# Check descriptives
library(psych);
library(knitr);
library(pastecs);
library(stargazer);
library(broom);
kable(describe(cbind(DDMtask.clean$rt, DDMtask.clean$log.rt)), digits = 2)
```

### Reminder for checking for normality

Normality?        | Positive           | Negative
------------      |----------------    |--------------
Skew              | too many low scores| too many high scores
Kurtosis          | heavy-tailed       | light-tailed

  + The further the value (skew or kurtosis) from zero, the more likely the distribution is not normal
  + Normally distributed scores will result in a straight diagonal line in the Q-Q plot. 

### Histogram of Reaction Time
```{r, echo = FALSE, results = "hide", message = FALSE, fig.width = 4, fig.height = 3}
# load packages
library(ggplot2);
```

``` {r, echo = FALSE, message = FALSE, fig.width = 4, fig.height = 3}
# create histogram object
hist.rt <- ggplot(DDMtask.clean, aes(rt)) + geom_histogram(aes(y = ..density..)) + labs(x = "Reaction Time", y = "Density") + theme_bw() + stat_function(fun = dnorm, args = list(mean = mean(DDMtask.clean$rt, na.rm = TRUE), sd = sd(DDMtask.clean$rt, na.rm = TRUE)), color = "red", size = 1);                                                                                                                                             
# display
print(hist.rt)
```

### Q-Q plot of Reaction Time
```{r, echo = FALSE, message = FALSE, fig.width = 4, fig.height = 3}
# q-q plot
qqplot.rt <- qplot(sample = DDMtask.clean$rt, stat = "qq") + theme_bw();
print(qqplot.rt)
```

### Histogram of logged Reaction Time
``` {r, echo = FALSE, message = FALSE, fig.width = 4, fig.height = 3}
# histogram
hist.log.rt <- ggplot(DDMtask.clean, aes(log.rt)) + geom_histogram(aes(y = ..density..)) + labs(x = "Reaction Time", y = "Density") + theme_bw() + stat_function(fun = dnorm, args = list(mean = mean(DDMtask.clean$log.rt, na.rm = TRUE), sd = sd(DDMtask.clean$log.rt, na.rm = TRUE)), color = "red", size = 1);                                                                                                                                             
print(hist.log.rt)
```

### Q-Q plot of logged Reaction Time
```{r, echo = FALSE, message = FALSE, fig.width = 4, fig.height = 3}
# q-q plot
qqplot.log.rt <- qplot(sample = DDMtask.clean$log.rt, stat = "qq") + theme_bw();
print(qqplot.log.rt)
```



**Check that multilevel model is needed:**
```{r, echo = FALSE, results = "asis", message = FALSE}
library(nlme);
library(pander)
library(xtable)
# check that we need a multilevel model
intercept.only <- gls(log.rt ~ 1, data = DDMtask.clean, method = "ML")
random.intercept.only <- lme(log.rt ~ 1, random = ~1|subject, data = DDMtask.clean, method = "ML", na.action = na.exclude, control = list(opt = "optim"))
# stargazer(intercept.only, random.intercept.only, type = "html", title = "Fixed vs. Random Intercept Only Models", digits = 2)
```

**Fixed Intercept only model**
```{r, echo = FALSE}
summary(intercept.only)
```

**Random Intercept only model**
```{r, echo = FALSE}
summary(random.intercept.only)
```


```{r, echo = FALSE, results = "asis", message = FALSE}
need.mlm <- anova(intercept.only, random.intercept.only)
# stargazer(need.mlm, summary = FALSE, type = "html", title = "Compare Fixed Intercept and Random Intercept Models", digits = 2, omit = c("call", "Model"))
```

**Compare Fixed Intercept vs. Random Intercept**
```{r, echo = FALSE}
need.mlm
```

  
**Fixed Effect Model**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
trial.fixed <- update(random.intercept.only, .~. + trial)
# stargazer(trial.fixed, type = "html", title = "Trial as Fixed Effect, No Predictors")
```

```{r, echo = FALSE}
summary(trial.fixed)
```


**Random Effect Model**
In this model, the intercept and the effect of trial vary across subjects.
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
trial.random <- update(trial.fixed, random = ~trial|subject)
# stargazer(trial.random, type = "html", title = "Trial as Random Effect, No Predictors")
```

```{r, echo = FALSE}
summary(trial.random)
```


**Full Effect Model with Predictor Deck Type**
In this model, the intercept and the effect of trial vary across subjects, AND the interaction between deck type and block is used as a fixed coefficient predictor.
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.one <- update(trial.random, .~. + deck.type)
# stargazer(model.one, type = "html", title = "Trial as Random Effect, Deck as Predictor")
# stargazer(model.one$coeff)
```

```{r, echo = FALSE}
summary(model.one)
```


**Full Effect Model with Predictors Deck Type & Block**
In this model, the intercept and the effect of trial vary across subjects, AND the interaction between deck type and block is used as a fixed coefficient predictor.
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.two <- update(model.one, .~. + block)
# stargazer(model.two, type = "html", title = "Trial as Random Effect, Deck & Block as Predictors")
```

```{r, echo = FALSE}
summary(model.two)
```

**Full Effect Model with Predictors Deck Type, Block, Card Number**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.three <- update(model.two, .~. + card.number)
```

```{r, echo = FALSE}
summary(model.three)
```


**Full Effect Model with Predictors Deck Type, Block, Card Number, Accept**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.four <- update(model.three, .~. + accept)
```

```{r, echo = FALSE}
summary(model.four)
```


**Full Effect Model with Predictors Deck Type, Block, Card Number, Accept, Confidence**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.five <- update(model.four, .~. + confidence)
```

```{r, echo = FALSE}
summary(model.five)
```


**Full Effect Model with Predictors Deck Type, Card Number, Accept, Confidence; Block is removed**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.six <- update(model.one, .~. + card.number + accept + confidence)
```

```{r, echo = FALSE}
summary(model.six)
```



#### The DV, Reaction Time, is not logged
**Full Effect Model with Predictors Deck Type, Card Number, Accept, Confidence; Block is removed**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.seven <- update(model.six, rt ~.)
```

```{r, echo = FALSE}
summary(model.seven)
```


**Full Effect Model with Predictors Deck Type, Card Number, Accept, Confidence, and DeckType:Confidence**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.eight <- update(model.seven, .~. + deck.type:confidence)
```

```{r, echo = FALSE}
summary(model.eight)
```

**Full Effect Model with Predictors Deck Type, Card Number, Accept, Confidence, and CardNumber:Confidence**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.nine <- update(model.seven, .~. + card.number:confidence)
```

```{r, echo = FALSE}
summary(model.nine)
```


**Full Effect Model with Predictors Deck Type, Card Number, Accept, Confidence, CardNumber:Confidence, and Accept:Confidence**
```{r, echo = FALSE, results = "asis", message = FALSE}
# update model to include time as a fixed effect
model.ten <- update(model.nine, .~. + accept:confidence)
```

```{r, echo = FALSE}
summary(model.ten)
plot(rt ~ trial, data = DDMtask.clean)
lines(predict(model.ten))
```


